soa

### 2024-10-19 inst/manuscript_figures.Rmd
gsea updates

# prefix <- "gsea_all"
# f <- file.path(results_dir, paste0(prefix, "_results.tsv"))
# data.table::fwrite(res, file=f, sep="\t", sep2=c("", " ", ""))
# f <- file.path(results_dir, paste0(prefix, "_results.xlsx"))
# write_xlsx(res, f)
```
### GSEA plots: helper functions
```{r helpers}
gsea_padj_cutoff <- 0.01
gsea_log2fc_cutoff <- 1

>>>>

# prefix <- "gsea_all"
# f <- file.path(results_dir, paste0(prefix, "_results.tsv"))
# data.table::fwrite(res, file=f, sep="\t", sep2=c("", " ", ""))
# f <- file.path(results_dir, paste0(prefix, "_results.xlsx"))
# write_xlsx(res, f)
```

### GSEA: all msigdb
```{r all_msigdb}
gsea_nperms <- 10000

# read msigdb data

f <- system.file("extdata", "msigdb.v2024.1.Hs.json", package = "stgeomx")
msigdb_gene_sets <- read_msigdb_json_file(f)

# determine gene universe
msigdb_genes <- intersect(gene_universe, msigdb_gene_sets$gene_symbol)

# filter genes not in our dataset
msigdb_gene_sets <- msigdb_gene_sets %>%
  dplyr::filter(gene_symbol %in% gene_universe) %>%
  distinct(gs_cat, gs_name, gene_symbol, .keep_all = TRUE)

# gene set categories
gs_hallmark <- dplyr::filter(msigdb_gene_sets, gs_cat == "H")
gs_hallmark <- split(x = gs_hallmark$gene_symbol, f = gs_hallmark$gs_name)
gs_gobp <- dplyr::filter(msigdb_gene_sets, gs_cat == "C5:GO:BP")
gs_gobp <- split(x = gs_gobp$gene_symbol, f = gs_gobp$gs_name)
gs_3ca <- dplyr::filter(msigdb_gene_sets, gs_cat == "C4:3CA")
gs_3ca <- split(x = gs_3ca$gene_symbol, f = gs_3ca$gs_name)

cp_cat <- c("C2:CP:BIOCARTA", "C2:CP:KEGG_MEDICUS", "C2:CP:PID", "C2:CP:REACTOME",
            "C2:CP:WIKIPATHWAYS")
gs_cp <- dplyr::filter(msigdb_gene_sets, gs_cat %in% cp_cat)
gs_cp <- split(x = gs_cp$gene_symbol, f = gs_cp$gs_name)
gs_tft <- dplyr::filter(msigdb_gene_sets, gs_cat == "C3:TFT:GTRD")
gs_tft <- split(x = gs_tft$gene_symbol, f = gs_tft$gs_name)

  res <- fgsea_run_batch(de, gs_gobp,
                       fgsea_eps=0, fgsea_nperms=gsea_nperms)
  res <- res %>% mutate(method = method)
  gsea <- bind_rows(gsea, res)

  res <- fgsea_run_batch(de, gs_3ca,
                       fgsea_eps=0, fgsea_nperms=gsea_nperms)
  res <- res %>% mutate(method = method)
  gsea <- bind_rows(gsea, res)

  res <- fgsea_run_batch(de, gs_cp,
                       fgsea_eps=0, fgsea_nperms=gsea_nperms)
  res <- res %>% mutate(method = method)
  gsea <- bind_rows(gsea, res)

  res <- fgsea_run_batch(de, gs_tft,
                       fgsea_eps=0, fgsea_nperms=gsea_nperms)
  res <- res %>% mutate(method = method)
  gsea <- bind_rows(gsea, res)

# write results for clusters
contrasts <- c("clusterPanc", "clusterBrain", "clusterLn", "clusterKidney", "clusterColon")
manuscript_pathways <- c(names(gs_hallmark), names(gs_gobp), names(gs_3ca), names(gs_cp), names(gs_tft))
res <- gsea_all %>% filter(contrast %in% contrasts,
                           pathway %in% manuscript_pathways)

prefix <- "gsea_all"
f <- file.path(results_dir, paste0(prefix, "_results.tsv"))
data.table::fwrite(res, file=f, sep="\t", sep2=c("", " ", ""))
f <- file.path(results_dir, paste0(prefix, "_results.xlsx"))
write_xlsx(res, f)
```

### GSEA plots: helper functions
```{r helpers}
gsea_padj_cutoff <- 0.01
gsea_log2fc_cutoff <- 1
